DOM Review Summary

### 2024-03-18 Analysis/ReviewSummary/ReviewSummary.qmd
Covariate changes plots
Plots for types of covs, attributes, changes in time, study size, etc

+-------------------+--------------------------------------------------+--------------------------------------+
| Field             | Description                                      | Possible values                      |
+===================+==================================================+======================================+
| Journal           | Journal which this article was published in      | *Any*                                |
+-------------------+--------------------------------------------------+--------------------------------------+
| Year              | Year of publication                              | *Any*                                |
+-------------------+--------------------------------------------------+--------------------------------------+
| Author country    | Nation of the first author's primary affiliation | *ISO 3166 country names*             |
+-------------------+--------------------------------------------------+--------------------------------------+
| Author categories | Categories of affiliations for *all* authors     | *Academic, Government, NGO, Private* |
+-------------------+--------------------------------------------------+--------------------------------------+

@@

+---------------------+-----------------------------------------------------------------------------------+
| Objective           | Description                                                                       |
+=====================+===================================================================================+
| Methods             | Did authors seek to demonstrate or test specific methods for occupancy modelling? |
+---------------------+-----------------------------------------------------------------------------------+
| Future Predictions  | Did authors make predictions beyond the study period?                             |
+---------------------+-----------------------------------------------------------------------------------+
| Spatial Predictions | Did authors make predictions at locations where surveys were not conducted?       |
+---------------------+-----------------------------------------------------------------------------------+
| Hypothesis testing  | Did authors seek to test specific hypotheses of species occupancy dynamics?       |
+---------------------+-----------------------------------------------------------------------------------+

@@

+-----------------+--------------------------------------------------------------------------------------------+-----------------------------------------------------------------------------------+
| Field           | Description                                                                                | Possible values                                                                   |
+=================+============================================================================================+===================================================================================+
| Taxa category   | Category of study taxa; could fit into multiple categories                                 | *Bird, Mammal, Invertebrate, Herptile, Fish, Threatened, Invasive, Other*         |
+-----------------+--------------------------------------------------------------------------------------------+-----------------------------------------------------------------------------------+
| Taxa number     | Number of individual species within the taxa for which models were fit                     | *Numeric*                                                                         |
+-----------------+--------------------------------------------------------------------------------------------+-----------------------------------------------------------------------------------+
| IUCN status     | IUCN status for the study species; NA if multi-species or if the study species is invasive | *LC, NT, EN, CR, DD, NA*                                                          |
+-----------------+--------------------------------------------------------------------------------------------+-----------------------------------------------------------------------------------+
| Threat category | Category of threats to the species; multiple per species permitted                         | *Habitat loss, Climate change, Exploitation, Invasives, Disease, Disaster, Other* |
+-----------------+--------------------------------------------------------------------------------------------+-----------------------------------------------------------------------------------+
| Country         | Country/countries where the study took place                                               | *ISO 3166 country names*                                                          |
+-----------------+--------------------------------------------------------------------------------------------+-----------------------------------------------------------------------------------+
| Continent       | Continent(s) where the study took place                                                    | *North America, South America, Europe, Asia, Africa, Oceania*                     |
+-----------------+--------------------------------------------------------------------------------------------+-----------------------------------------------------------------------------------+
| Realm           | Biogeographic realm where the study took place                                             | *Nearctic, Palearctic, Afrotropical, Indomalayan, Australasian, Neotropical*      |
+-----------------+--------------------------------------------------------------------------------------------+-----------------------------------------------------------------------------------+
| Size            | Size of the study area extent in square kilometres                                         | *\<1km, 1-10, 10-100, 100-1000, 10000-100000, 100000-1000000, 1000000+*           |
+-----------------+--------------------------------------------------------------------------------------------+-----------------------------------------------------------------------------------+

@@

+---------------------+--------------------------------------------------------------------------+----------------------------------------------------------------------------------------------+
| Field               | Description                                                              | Possible values                                                                              |
+=====================+==========================================================================+==============================================================================================+
| Duration            | Length of study in months                                                | *Numeric*                                                                                    |
+---------------------+--------------------------------------------------------------------------+----------------------------------------------------------------------------------------------+
| Capture method      | Method of recording detections; multiple categories possible             | *Audiovisual, Search, Trapping, Bioacoustic, Camera, Citizen, Incidental*                    |
+---------------------+--------------------------------------------------------------------------+----------------------------------------------------------------------------------------------+
| Primary occasions   | Number of primary occasions                                              | *Numeric*                                                                                    |
+---------------------+--------------------------------------------------------------------------+----------------------------------------------------------------------------------------------+
| Secondary occasions | Number of secondary occasions, presented as a range where not consistent | *Numeric range*                                                                              |
+---------------------+--------------------------------------------------------------------------+----------------------------------------------------------------------------------------------+
| Secondary type      | Type of secondary occasion                                               | *Revisits, Spatial replicates, Same-visit replicates, multiple observers, time to detection* |
+---------------------+--------------------------------------------------------------------------+----------------------------------------------------------------------------------------------+
| Type of site        | How a 'site' was defined                                                 | *Point location, Grid cell, Irregular area, Transect*                                        |
+---------------------+--------------------------------------------------------------------------+----------------------------------------------------------------------------------------------+
| Number of sites     | Number of individual sites                                               | Numeric                                                                                      |
+---------------------+--------------------------------------------------------------------------+----------------------------------------------------------------------------------------------+
| Sampling method     | How sites were selected                                                  | *Random, Stratified, Grid, Preferential, Available, Exhaustive, Uniform, Transect*           |
+---------------------+--------------------------------------------------------------------------+----------------------------------------------------------------------------------------------+

@@

covariates_clean <- review_Covariates |>
  filter(Status == "Complete") |>
  mutate(ID_clean = str_extract(ID, "[:digit:]+")) |>
  mutate(Covariates = str_replace(Covariates, "\\.", "\\,")) |>
  separate_rows(Covariates, sep = ", ") |>
  mutate(Covariates = str_to_upper(Covariates)) |>
  separate_wider_delim(cols = Covariates, delim = "_",
                       names = c("Type", 
                                 "Variation",
                                 "Measurement",
                                 "Complexity"),
                       too_few = "align_start", 
                       too_many = "drop") |>
  mutate(ParameterCategory = case_when(Parameter %in% c("InitialOccupancy",
                                                        "InitialOccupancy_MS",
                                                        "Occupancy",
                                                        "Occupancy_MS") ~ "Psi",
                                       Parameter %in% c("Colonisation",
                                                        "Colonisation_MS") ~ "Gamma",
                                       Parameter %in% c("Extinction",
                                                        "Extinction_MS",
                                                        "Persistence",
                                                        "Persistence_MS") ~ "Eps",
                                       Parameter %in% c("Detection",
                                                        "Detection_FP",
                                                        "Detection_FP_MS",
                                                        "Detection_MS", 
                                                        "Detection_Other") ~ "p",
                                       .default = "Other"),
         Type = case_when(Type %in% c("E-HTT", " E-HTT") ~ "E-HTT",
                          Type %in% c("E-DIS", "DIS") ~ "E-DIS",
                          Type %in% c("ANTH", "E-ANTH") ~ "E-ANTH",
                          Type %in% c("SP", "Spec") ~ "SP",
                          Type %in% c("NONE", "NA") ~ "NONE",
                          .default = Type),
         Variation = case_when(Variation %in% c("SY", "YS") ~ "Site-Year",
                               Variation == "S" ~ "Site",
                               Variation == "YSS" ~ "Site-Year-Survey",
                               Variation == "Y" ~ "Year",
                               Variation == "SP" ~ "Species",
                               Variation == "U" ~ "Unclear"),
         Measurement = case_when(Measurement == "I" ~ "Intrinsic",
                                 Measurement == "O" ~ "Observed",
                                 Measurement == "R" ~ "Remote",
                                 Measurement == "U" ~ "Unclear"),
         Complexity = case_when(Complexity %in% c("N", "NH", "N,") ~ "Linear",
                                Complexity == "Q" ~ "Quadratic",
                                Complexity == "I" ~ "Interaction",
                                Complexity == "B" ~ "Quadratic + Interaction",
                                Complexity == "O" ~ "Other"))
```

@@

@@

## Frequency of covariate features

@@

@@

@@

+---------------+-----------------------------------------------------------+------------------------------------------------------------------------------------------------+
| Field         | Description                                               | Possible values                                                                                |
+===============+===========================================================+================================================================================================+
| Bayesian      | Whether the models fit were Bayesian                      | *Logical*                                                                                      |
+---------------+-----------------------------------------------------------+------------------------------------------------------------------------------------------------+
| Prefiltering  | Which, if any, prefiltering method was used               | *Expert, Simple, Other*                                                                        |
+---------------+-----------------------------------------------------------+------------------------------------------------------------------------------------------------+
| Selection     | Which, if any, selection method was used                  | *Single a priori, Multiple a priori, Exhaustive, Exhaustive by parameter, Forwards, Backwards* |
+---------------+-----------------------------------------------------------+------------------------------------------------------------------------------------------------+
| Evaluation    | How models were compared and evaluated                    | *AIC, AUC, DiC, Model-averaging, Bootstrapping*                                                |
+---------------+-----------------------------------------------------------+------------------------------------------------------------------------------------------------+
| Multi-species | Whether this was a multi-species model                    | *Logical*                                                                                      |
+---------------+-----------------------------------------------------------+------------------------------------------------------------------------------------------------+
| Connectivity  | Whether site connectivity was incorporated into the model | *Logical*                                                                                      |
+---------------+-----------------------------------------------------------+------------------------------------------------------------------------------------------------+

>>>>

| Field             | Description                                      | Possible values                      |
|-------------------|--------------------------------------------------|--------------------------------------|
| Journal           | Journal which this article was published in      | *Any*                                |
| Year              | Year of publication                              | *Any*                                |
| Author country    | Nation of the first author's primary affiliation | *ISO 3166 country names*             |
| Author categories | Categories of affiliations for *all* authors     | *Academic, Government, NGO, Private* |

@@

| Objective           | Description                                                                       |
|---------------------|-----------------------------------------------------------------------------------|
| Methods             | Did authors seek to demonstrate or test specific methods for occupancy modelling? |
| Future Predictions  | Did authors make predictions beyond the study period?                             |
| Spatial Predictions | Did authors make predictions at locations where surveys were not conducted?       |
| Hypothesis testing  | Did authors seek to test specific hypotheses of species occupancy dynamics?       |

@@

| Field           | Description                                                                                | Possible values                                                                   |
|-----------------|--------------------------------------------------------------------------------------------|-----------------------------------------------------------------------------------|
| Taxa category   | Category of study taxa; could fit into multiple categories                                 | *Bird, Mammal, Invertebrate, Herptile, Fish, Threatened, Invasive, Other*         |
| Taxa number     | Number of individual species within the taxa for which models were fit                     | *Numeric*                                                                         |
| IUCN status     | IUCN status for the study species; NA if multi-species or if the study species is invasive | *LC, NT, EN, CR, DD, NA*                                                          |
| Threat category | Category of threats to the species; multiple per species permitted                         | *Habitat loss, Climate change, Exploitation, Invasives, Disease, Disaster, Other* |
| Country         | Country/countries where the study took place                                               | *ISO 3166 country names*                                                          |
| Continent       | Continent(s) where the study took place                                                    | *North America, South America, Europe, Asia, Africa, Oceania*                     |
| Realm           | Biogeographic realm where the study took place                                             | *Nearctic, Palearctic, Afrotropical, Indomalayan, Australasian, Neotropical*      |
| Size            | Size of the study area extent in square kilometres                                         | *\<1km, 1-10, 10-100, 100-1000, 10000-100000, 100000-1000000, 1000000+*           |

@@

| Field               | Description                                                              | Possible values                                                                              |
|---------------------|--------------------------------------------------------------------------|----------------------------------------------------------------------------------------------|
| Duration            | Length of study in months                                                | *Numeric*                                                                                    |
| Capture method      | Method of recording detections; multiple categories possible             | *Audiovisual, Search, Trapping, Bioacoustic, Camera, Citizen, Incidental*                    |
| Primary occasions   | Number of primary occasions                                              | *Numeric*                                                                                    |
| Secondary occasions | Number of secondary occasions, presented as a range where not consistent | *Numeric range*                                                                              |
| Secondary type      | Type of secondary occasion                                               | *Revisits, Spatial replicates, Same-visit replicates, multiple observers, time to detection* |
| Type of site        | How a 'site' was defined                                                 | *Point location, Grid cell, Irregular area, Transect*                                        |
| Number of sites     | Number of individual sites                                               | Numeric                                                                                      |
| Sampling method     | How sites were selected                                                  | *Random, Stratified, Grid, Preferential, Available, Exhaustive, Uniform, Transect*           |

@@

covariates_clean <- review_Covariates |>
  filter(Status == "Complete") |>
  mutate(ID_clean = str_extract(ID, "[:digit:]+")) |>
  mutate(Covariates = str_replace(Covariates, "\\.", "\\,")) |>
  separate_rows(Covariates, sep = ", ") |>
  mutate(Covariates = str_to_upper(Covariates)) |>
  separate_wider_delim(cols = Covariates, delim = "_",
                       names = c("Type", 
                                 "Variation",
                                 "Measurement",
                                 "Complexity"),
                       too_few = "align_start", 
                       too_many = "drop") |>
  mutate(ParameterCategory = case_when(Parameter %in% c("InitialOccupancy",
                                                        "InitialOccupancy_MS",
                                                        "Occupancy",
                                                        "Occupancy_MS") ~ "Psi",
                                       Parameter %in% c("Colonisation",
                                                        "Colonisation_MS") ~ "Gamma",
                                       Parameter %in% c("Extinction",
                                                        "Extinction_MS",
                                                        "Persistence",
                                                        "Persistence_MS") ~ "Eps",
                                       Parameter %in% c("Detection",
                                                        "Detection_FP",
                                                        "Detection_FP_MS",
                                                        "Detection_MS", 
                                                        "Detection_Other") ~ "p",
                                       .default = "Other"),
         Type = case_when(Type %in% c("E-HTT", " E-HTT") ~ "E-HTT",
                          Type %in% c("E-DIS", "DIS") ~ "E-DIS",
                          Type %in% c("ANTH", "E-ANTH") ~ "E-ANTH",
                          Type %in% c("SP", "SPEC") ~ "SP",
                          Type %in% c("INT", " INT") ~ "INT",
                          Type %in% c("NONE", "NA") ~ "NONE",
                          .default = Type),
         Variation = case_when(Variation %in% c("SY", "YS") ~ "Site-Year",
                               Variation == "S" ~ "Site",
                               Variation == "YSS" ~ "Site-Year-Survey",
                               Variation == "Y" ~ "Year",
                               Variation == "SP" ~ "Species",
                               Variation == "U" ~ "Unclear"),
         Measurement = case_when(Measurement == "I" ~ "Intrinsic",
                                 Measurement == "O" ~ "Observed",
                                 Measurement == "R" ~ "Remote",
                                 Measurement == "U" ~ "Unclear"),
         Complexity = case_when(Complexity %in% c("N", "NH", "N,") ~ "Linear",
                                Complexity == "Q" ~ "Quadratic",
                                Complexity == "I" ~ "Interaction",
                                Complexity == "B" ~ "Quadratic + Interaction",
                                Complexity == "O" ~ "Other"))
```

@@

::: column-page
::: panel-tabset
## Overall

```{r covariate_categoriesOverall}
#| echo: false

covariates_clean |>
  mutate(Total = length(unique(ID))) |>
  select(ID, Type, Total) |>
  distinct() |>
  summarise(Articles = n(),
            .by = c(Type, Total)) |>
  mutate(Percentage = round(Articles/Total, 2)*100) |>
  arrange(-Percentage) |>
  filter(!Type %in% c("UNSPECIFIED", "UNCLEAR", "NONE", NA, "OTHER")) |>
  mutate(Type = case_when(Type == "E-ANTH" ~ "Anthropogenic",
                          Type == "E-CLIM" ~ "Climate/Weather",
                          Type == "E-DIS" ~ "Disaster/Disturbance",
                          Type == "E-HTAQ" ~ "Habitat (Aquatic)",
                          Type == "E-HTT" ~ "Habitat (Terrestrial)",
                          Type == "E-HYD" ~ "Hydrologic",
                          Type == "E-SEAS" ~ "Phenologic",
                          Type == "E-TOPO" ~ "Topographic",
                          Type == "CONN" ~ "Connectivity",
                          Type == "GEOM" ~ "Site geometry",
                          Type == "INT" ~ "Biotic interaction",
                          Type == "SEAS" ~ "Primary occasion",
                          Type == "SITE" ~ "Site",
                          Type == "SP" ~ "Species",
                          Type == "SURV" ~ "Survey detail")) |>
  mutate(Type_cat = case_when(Type %in% c("Habitat (Terrestrial)",
                                          "Habitat (Aquatic)",
                                          "Hydrologic", 
                                          "Climate/Weather",
                                          "Disaster/Disturbance",
                                          "Topographic",
                                          "Phenologic",
                                          "Anthropogenic",
                                          "Biotic interaction") ~ "Env",
                              .default = "Other")) |>
  mutate(Type = fct(Type,
                    levels = c("Habitat (Terrestrial)", "Habitat (Aquatic)",
                               "Hydrologic", "Climate/Weather",
                               "Disaster/Disturbance", "Topographic",
                               "Phenologic", "Anthropogenic",
                               "Biotic interaction", "Primary occasion",
                               "Survey detail", "Species", 
                               "Site", "Site geometry", "Connectivity"))) |>
  select(Type, Type_cat, Percentage) |>

  ggplot() +
  geom_col(aes(x = Type, y = Percentage, fill = Type_cat)) +
  labs(title = "Most frequent covariate types (All parameters)",
       x = "", y = "Percentage of models (any parameter)") +
  scale_fill_manual(values = c("darkslategrey", "darkorange3")) +
  theme(axis.text.x = element_text(angle = 90, 
                                   vjust = 0.5, hjust = 1),
        legend.position = "none")
```

## Initial Occupancy

```{r Covariate_categoriesPsi}
#| echo: false

covariates_clean |>
  mutate(Total = length(unique(ID))) |>
  filter(ParameterCategory == "Psi") |>
  select(ID, Type, Total) |>
  distinct() |>
  summarise(Articles = n(),
            .by = c(Type, Total)) |>
  mutate(Percentage = round(Articles/Total, 2)*100) |>
  arrange(-Percentage) |>
  filter(!Type %in% c("UNSPECIFIED", "UNCLEAR", "NONE", NA, "SURV", "SEAS", "OTHER")) |>
  mutate(Type = case_when(Type == "E-ANTH" ~ "Anthropogenic",
                          Type == "E-CLIM" ~ "Climate/Weather",
                          Type == "E-DIS" ~ "Disaster/Disturbance",
                          Type == "E-HTAQ" ~ "Habitat (Aquatic)",
                          Type == "E-HTT" ~ "Habitat (Terrestrial)",
                          Type == "E-HYD" ~ "Hydrologic",
                          Type == "E-SEAS" ~ "Phenologic",
                          Type == "E-TOPO" ~ "Topographic",
                          Type == "CONN" ~ "Connectivity",
                          Type == "GEOM" ~ "Site geometry",
                          Type == "INT" ~ "Biotic interaction",
                          Type == "SEAS" ~ "Primary occasion",
                          Type == "SITE" ~ "Site",
                          Type == "SP" ~ "Species",
                          Type == "SURV" ~ "Survey detail",
                          Type == "OTHER" ~ "Other")) |>  
  mutate(Type_cat = case_when(Type %in% c("Habitat (Terrestrial)",
                                          "Habitat (Aquatic)",
                                          "Hydrologic", 
                                          "Climate/Weather",
                                          "Disaster/Disturbance",
                                          "Topographic",
                                          "Phenologic",
                                          "Anthropogenic",
                                          "Biotic interaction") ~ "Env",
                              .default = "Other")) |>
  mutate(Type = fct(Type,
                    levels = c("Habitat (Terrestrial)", "Habitat (Aquatic)",
                               "Hydrologic", "Climate/Weather",
                               "Disaster/Disturbance", "Topographic",
                               "Phenologic", "Anthropogenic",
                               "Biotic interaction", "Species", 
                               "Site", "Site geometry", "Connectivity"))) |>
  select(Type, Type_cat, Percentage) |>

  ggplot() +
  geom_col(aes(x = Type, y = Percentage, fill = Type_cat)) +
  labs(title = "Most frequent covariate types (Initial Occupancy)",
       x = "", y = "Percentage of models") +
  scale_fill_manual(values = c("darkslategrey", "darkorange3")) +
  theme(axis.text.x = element_text(angle = 90, 
                                   vjust = 0.5, hjust = 1),
        legend.position = "none")
```

## Colonisation

```{r covariates_catrgoriesGamma}
#| echo: false

covariates_clean |>
  mutate(Total = length(unique(ID))) |>
  filter(ParameterCategory == "Gamma") |>
  select(ID, Type, Total) |>
  distinct() |>
  summarise(Articles = n(),
            .by = c(Type, Total)) |>
  mutate(Percentage = round(Articles/Total, 2)*100) |>
  arrange(-Percentage) |>
  filter(!Type %in% c("UNSPECIFIED", "UNCLEAR", "NONE", NA, "SURV", "OTHER")) |>
  mutate(Type = case_when(Type == "E-ANTH" ~ "Anthropogenic",
                          Type == "E-CLIM" ~ "Climate/Weather",
                          Type == "E-DIS" ~ "Disaster/Disturbance",
                          Type == "E-HTAQ" ~ "Habitat (Aquatic)",
                          Type == "E-HTT" ~ "Habitat (Terrestrial)",
                          Type == "E-HYD" ~ "Hydrologic",
                          Type == "E-SEAS" ~ "Phenologic",
                          Type == "E-TOPO" ~ "Topographic",
                          Type == "CONN" ~ "Connectivity",
                          Type == "GEOM" ~ "Site geometry",
                          Type == "INT" ~ "Biotic interaction",
                          Type == "SEAS" ~ "Primary occasion",
                          Type == "SITE" ~ "Site",
                          Type == "SP" ~ "Species",
                          Type == "SURV" ~ "Survey detail",
                          Type == "OTHER" ~ "Other")) |>  
  mutate(Type_cat = case_when(Type %in% c("Habitat (Terrestrial)",
                                          "Habitat (Aquatic)",
                                          "Hydrologic", 
                                          "Climate/Weather",
                                          "Disaster/Disturbance",
                                          "Topographic",
                                          "Phenologic",
                                          "Anthropogenic",
                                          "Biotic interaction") ~ "Env",
                              .default = "Other")) |>
  mutate(Type = fct(Type,
                    levels = c("Habitat (Terrestrial)", "Habitat (Aquatic)",
                               "Hydrologic", "Climate/Weather",
                               "Disaster/Disturbance", "Topographic",
                               "Phenologic", "Anthropogenic",
                               "Biotic interaction", "Species",
                               "Primary occasion", "Site",
                               "Site geometry", "Connectivity"))) |>
  select(Type, Type_cat, Percentage) |>

  ggplot() +
  geom_col(aes(x = Type, y = Percentage, fill = Type_cat)) +
  labs(title = "Most frequent covariate types (Colonisation)",
       x = "", y = "Percentage of models") +
  scale_fill_manual(values = c("darkslategrey", "darkorange3")) +
  theme(axis.text.x = element_text(angle = 90, 
                                   vjust = 0.5, hjust = 1),
        legend.position = "none")
```

## Extinction

```{r covariates_categoryEps}
#| echo: false

covariates_clean |>
  mutate(Total = length(unique(ID))) |>
  filter(ParameterCategory == "Eps") |>
  select(ID, Type, Total) |>
  distinct() |>
  summarise(Articles = n(),
            .by = c(Type, Total)) |>
  mutate(Percentage = round(Articles/Total, 2)*100) |>
  arrange(-Percentage) |>
  filter(!Type %in% c("UNSPECIFIED", "UNCLEAR", "NONE", NA, "SURV", "OTHER")) |>
  mutate(Type = case_when(Type == "E-ANTH" ~ "Anthropogenic",
                          Type == "E-CLIM" ~ "Climate/Weather",
                          Type == "E-DIS" ~ "Disaster/Disturbance",
                          Type == "E-HTAQ" ~ "Habitat (Aquatic)",
                          Type == "E-HTT" ~ "Habitat (Terrestrial)",
                          Type == "E-HYD" ~ "Hydrologic",
                          Type == "E-SEAS" ~ "Phenologic",
                          Type == "E-TOPO" ~ "Topographic",
                          Type == "CONN" ~ "Connectivity",
                          Type == "GEOM" ~ "Site geometry",
                          Type == "INT" ~ "Biotic interaction",
                          Type == "SEAS" ~ "Primary occasion",
                          Type == "SITE" ~ "Site",
                          Type == "SP" ~ "Species",
                          Type == "SURV" ~ "Survey detail",
                          Type == "OTHER" ~ "Other")) |>  
  mutate(Type_cat = case_when(Type %in% c("Habitat (Terrestrial)",
                                          "Habitat (Aquatic)",
                                          "Hydrologic", 
                                          "Climate/Weather",
                                          "Disaster/Disturbance",
                                          "Topographic",
                                          "Phenologic",
                                          "Anthropogenic",
                                          "Biotic interaction") ~ "Env",
                              .default = "Other")) |>
  mutate(Type = fct(Type,
                    levels = c("Habitat (Terrestrial)", "Habitat (Aquatic)",
                               "Hydrologic", "Climate/Weather",
                               "Disaster/Disturbance", "Topographic",
                               "Phenologic", "Anthropogenic",
                               "Biotic interaction", "Species",
                               "Primary occasion", "Site",
                               "Site geometry", "Connectivity"))) |>
  select(Type, Type_cat, Percentage) |>

  ggplot() +
  geom_col(aes(x = Type, y = Percentage, fill = Type_cat)) +
  labs(title = "Most frequent covariate types (Extinction)",
       x = "", y = "Percentage of models") +
  scale_fill_manual(values = c("darkslategrey", "darkorange3")) +
  theme(axis.text.x = element_text(angle = 90, 
                                   vjust = 0.5, hjust = 1),
        legend.position = "none")

## Detection

```{r covariate_categoriesDet}
#| echo: false

covariates_clean |>
  mutate(Total = length(unique(ID))) |>
  filter(ParameterCategory == "p") |>
  select(ID, Type, Total) |>
  distinct() |>
  summarise(Articles = n(),
            .by = c(Type, Total)) |>
  mutate(Percentage = round(Articles/Total, 2)*100) |>
  arrange(-Percentage) |>
  filter(!Type %in% c("UNSPECIFIED", "UNCLEAR", "NONE", "OTHER", NA)) |>
  mutate(Type = case_when(Type == "E-ANTH" ~ "Anthropogenic",
                          Type == "E-CLIM" ~ "Climate/Weather",
                          Type == "E-DIS" ~ "Disaster/Disturbance",
                          Type == "E-HTAQ" ~ "Habitat (Aquatic)",
                          Type == "E-HTT" ~ "Habitat (Terrestrial)",
                          Type == "E-HYD" ~ "Hydrologic",
                          Type == "E-SEAS" ~ "Phenologic",
                          Type == "E-TOPO" ~ "Topographic",
                          Type == "CONN" ~ "Connectivity",
                          Type == "GEOM" ~ "Site geometry",
                          Type == "INT" ~ "Biotic interaction",
                          Type == "SEAS" ~ "Primary occasion",
                          Type == "SITE" ~ "Site",
                          Type == "SP" ~ "Species",
                          Type == "SURV" ~ "Survey detail",
                          Type == "OTHER" ~ "Other")) |>  
  mutate(Type_cat = case_when(Type %in% c("Habitat (Terrestrial)",
                                          "Habitat (Aquatic)",
                                          "Hydrologic", 
                                          "Climate/Weather",
                                          "Disaster/Disturbance",
                                          "Topographic",
                                          "Phenologic",
                                          "Anthropogenic",
                                          "Biotic interaction") ~ "Env",
                              .default = "Other")) |>
  mutate(Type = fct(Type,
                    levels = c("Habitat (Terrestrial)", "Habitat (Aquatic)",
                               "Hydrologic", "Climate/Weather",
                               "Disaster/Disturbance", "Topographic",
                               "Phenologic", "Anthropogenic",
                               "Biotic interaction", "Species",
                               "Primary occasion", "Site", "Survey detail",
                               "Site geometry", "Connectivity"))) |>
  select(Type, Type_cat, Percentage) |>

  ggplot() +
  geom_col(aes(x = Type, y = Percentage, fill = Type_cat)) +
  labs(title = "Most frequent covariate types (Detection)",
       x = "", y = "Percentage of models") +
  scale_fill_manual(values = c("darkslategrey", "darkorange3")) +
  theme(axis.text.x = element_text(angle = 90, 
                                   vjust = 0.5, hjust = 1),
        legend.position = "none")

```
:::
:::

@@

Aspects of each covariate are presented here as percentages; e.g., the frequency that a class of covariate had a quadratic term, or the frequency at which it was measured via remote sensing.

::: column-page
```{r Covariates_Features table}
#| echo: false

covariates_clean |> 
  mutate(Overall_Total = length(unique(ID))) |>
  mutate(Meas_Observed = case_when(Measurement == "Observed" ~ 1,
                                   .default = 0),
         Meas_Intrinsic = case_when(Measurement == "Intrinsic" ~ 1,
                                    .default = 0),
         Meas_Remote = case_when(Measurement == "Remote" ~ 1,
                                 .default = 0),
         Dynamic = case_when(Variation %in% c("Year",
                                              "Year-Site", 
                                              "Year-Site-Survey") ~ 1,
                             .default =0),
         Linear = case_when(Complexity == "Linear" ~ 1,
                               .default = 0),
         Quadratic = case_when(Complexity %in% c("Quadratic + Interaction",
                                                 "Quadratic") ~ 1,
                               .default = 0),
         Interaction = case_when(Complexity %in% c("Quadratic + Interaction",
                                                   "Interaction") ~ 1,
                               .default = 0)) |>
  summarise(Meas_Observed = sum(Meas_Observed)/n(),
            Meas_Intrinsic = sum(Meas_Intrinsic)/n(),
            Meas_Remote = sum(Meas_Remote)/n(),
            Dynamic = sum(Dynamic)/n(),
            Linear = sum(Linear)/n(),
            Quadratic = sum(Quadratic)/n(),
            Interaction = sum(Interaction)/n(),
            .by = Type) |>
  mutate(across(where(is.numeric), ~round(.x, 2))) |>
  mutate(Type = case_when(Type == "E-ANTH" ~ "Anthropogenic",
                          Type == "E-CLIM" ~ "Climate/Weather",
                          Type == "E-DIS" ~ "Disaster/Disturbance",
                          Type == "E-HTAQ" ~ "Habitat (Aquatic)",
                          Type == "E-HTT" ~ "Habitat (Terrestrial)",
                          Type == "E-HYD" ~ "Hydrologic",
                          Type == "E-SEAS" ~ "Phenologic",
                          Type == "E-TOPO" ~ "Topographic",
                          Type == "CONN" ~ "Connectivity",
                          Type == "GEOM" ~ "Site geometry",
                          Type == "INT" ~ "Biotic interaction",
                          Type == "SEAS" ~ "Primary occasion",
                          Type == "SITE" ~ "Site",
                          Type == "SP" ~ "Species",
                          Type == "SURV" ~ "Survey detail",
                          Type == "OTHER" ~ "Other"))|>
  filter(!is.na(Type) & Type != "Other") |>
  reactable()

## Complexity

21% of articles considered at least one quadratic term; 26% considered at least one interaction term.

```{r covariates_complexity}
#| echo: false

covariates_clean |>
  mutate(Total = length(unique(ID))) |>
  mutate(Quadratic = case_when(Complexity %in% c("Quadratic + Interaction",
                                                 "Quadratic") ~ 1,
                               .default = 0),
         Interaction = case_when(Complexity %in% c("Quadratic + Interaction",
                                                   "Interaction") ~ 1,
                                 .default = 0)) |>
  select(ID, Total, Quadratic, Interaction) |>
  distinct() |>
  summarise(Quadratic = round(sum(Quadratic)/mean(Total), 2),
            Interaction = round(sum(Interaction)/mean(Total),2)) |>
  kable()
```

@@

There is no apparent trend in the quantity of covariates over time.

```{r covariates_quantityYearly}
#| echo: false
#| message: false

covariates_clean |>
  filter(Type != "NONE") |>
  summarise(Covariates = n(), .by = ID_clean) |>
  mutate(ID_clean = as.numeric(ID_clean)) |>
  left_join(select(studyDetails_clean, Year, ID), 
            by = join_by(ID_clean == ID)) |>
  filter(Covariates < 60) |>

  ggplot(aes(x = Year, y = Covariates)) +
  geom_point(colour = "darkslategrey", alpha = 0.5) +
  geom_smooth(method = "lm", colour = "darkslategrey") +
  labs(title = "Quantities of covariates over time",
       x = "Year", y = "Total number of covariates")
```

@@

```{r Covariates_studysize}
#| echo: false
#| warning: false
#| message: false

covariates_count <- covariates_clean |>
  filter(Type != "NONE") |>
  summarise(Covariates = n(), .by = ID_clean) |>
  mutate(ID_clean = as.numeric(ID_clean))

dataCollection_clean |>
  select(ID_clean, Duration, 
         Primary_Max, Secondary_Max,
         SiteNumber_Max) |>
  mutate(ID_clean = as.numeric(ID_clean)) |>
  rowwise() |>
  mutate(TotalOccasions = Primary_Max*Secondary_Max) |>
  left_join(select(studyDetails_clean, ID, YearStrata), by = join_by(ID_clean == ID)) |>
  filter(!is.na(YearStrata)) |>
  pivot_longer(cols = c(Duration, TotalOccasions, SiteNumber_Max),
               names_to = "Variable", values_to = "Quantity") |>
  left_join(covariates_count, by = "ID_clean") |>
  mutate(Variable = case_when(Variable == "Duration" ~ "Study duration (months)",
                              Variable == "SiteNumber_Max" ~ "Number of sites",
                              Variable == "TotalOccasions" ~ "Total survey occasions (Primary x Secondary)")) |>
  filter(Covariates < 60) |>

  ggplot(aes(x = Quantity, y = Covariates, colour = Variable)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE, alpha = 0.7) +
  scale_x_continuous(transform = "log10") +
  scale_colour_manual(values = c("darkslategrey", "cyan4", "darkorchid4")) +
  facet_wrap(vars(Variable), ncol = 1, scales = "free_x") +
  theme(legend.position = "none") +
  labs(title = "Covariate count by study size",
       x = "Study size (log scales)", y = "Number of covariates")
```

## Covariate count by objective

```{r Covariates_objective}
#| echo: false

objectives_Clean |>
  select(ID, Objective, Stated) |>
  filter(Stated == TRUE) |>
  left_join(covariates_count, by = join_by(ID == ID_clean)) |>
  filter(Covariates < 60) |>

  ggplot() +
  geom_boxplot(aes(x = Objective, y = Covariates, colour = Objective)) +
  theme(legend.position = "none") +
  labs(title = "Total covariate count by objective",
       x = "", y = "Number of covariates") + 
  scale_fill_manual(values = c("darkorchid", "deepskyblue3",
                               "coral2", "goldenrod", "gray60"))
```

@@

| Field         | Description                                               | Possible values                                                                                |
|---------------|-----------------------------------------------------------|------------------------------------------------------------------------------------------------|
| Bayesian      | Whether the models fit were Bayesian                      | *Logical*                                                                                      |
| Prefiltering  | Which, if any, prefiltering method was used               | *Expert, Simple, Other*                                                                        |
| Selection     | Which, if any, selection method was used                  | *Single a priori, Multiple a priori, Exhaustive, Exhaustive by parameter, Forwards, Backwards* |
| Evaluation    | How models were compared and evaluated                    | *AIC, AUC, DiC, Model-averaging, Bootstrapping*                                                |
| Multi-species | Whether this was a multi-species model                    | *Logical*                                                                                      |
| Connectivity  | Whether site connectivity was incorporated into the model | *Logical*                                                                                      |
